An efficient composition of bidirectional programs by tupling and lazy updates

### 2020-02-21 paper/4.xpg.tex
modified the beginning of 4.4

When evaluating $\pginline{bx_1 \circ bx_2}$, there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$. If a given program is a left associative composition, the number of $pg$ calls will be exponential. Therefore, the runtime inefficiency is inevitable. To solve that, we introduce a new function, $cpg$, accumulates updates on the source and the view. $\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values same as $pg$. The output of this function is a 4-tuple $(ks, kv, s, v)$. To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:\\ $fst = \lambda (x_1,x_2). x_1 , \ snd = \lambda (x_1,x_2). x_2 , \ con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$

@@

In the places where third and/or fourth argument ($s$ and $v$) are updated by applications, the computations are also accumulated in $ks$ and/or $kv$.
%Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
Thanks to these accumulations, there are only two $cpg$ calls in $\cpginline{bx_1 \circ bx_2}$. The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, dummy)$ where $s$ and $ks$ are corresponding to the source and the update over source. Since there is no real view here, we need a dummy same as $pg$. Then the continuation updating on this dummy should be initiated as an identity function. The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$. In the next assignment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input. It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$. After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling recursively one more time like in $\pginline{bx_1 \circ bx_2}$.

@@

In general, the beginning of a continuation should be an identity function. However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source needs to be initiated as the constant function from that source. This constant function helps to retain the discarded things in the source.

@@

The problem for $cpg$ lies in redundant computations during the evaluation. To prevent such redundant computations from occurring, we introduce an extension named $kpg$. While $cpg$ evaluates values eagerly, $kpg$ does the opposite. Every value is evaluated lazily in a computation of $kpg$. The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, lks, lkv, s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $lks$ and $lkv$ are used for lazy evaluation of actual values. The output of this function is also a 6-tuple $(ks, kv, lks, lkv, s, v)$.\\
Suppose that we have a source $s_0$ and a view $v_0$. The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:

@@

The beginning of accumulative functions $lks$ and $lkv$ are set as identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$. The relationship among $kpg$, $cpg$ and $pg$ can be shown as follows:

@@

\subsection{Combination of pg and kpg: xpg}
\textcolor{red}{TODO-D3/D22: change explanation clearer.}
The most important thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that are used to produce new source instead of using more recursive call. This is reflected in the first call $\kpginline{bx_1}$ and the last application in $\kpginline{bx_1 \circ bx_2}$. The second call $\kpginline{bx_2}$ just play a syntactic role while it requires many parameters. We can completely reduce this number to 2, including one for the source and one for the view as in a $pg$ call. This idea is realized by function $xpg$ as follows:

>>>>

When evaluating $\pginline{bx_1 \circ bx_2}$, there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$. If a given program is a left associative composition, the number of $pg$ calls will be exponential. Therefore, the runtime inefficiency of $pg$ for left associative BX programs is inevitable. To solve that, we introduce a new function, $cpg$, accumulates updates on the source and the view. $\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values same as $pg$. The output of this function is a 4-tuple $(ks, kv, s, v)$. To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:\\ $fst = \lambda (x_1,x_2). x_1 , \ snd = \lambda (x_1,x_2). x_2 , \ con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$

@@

In the places where third and/or fourth argument ($s$ and $v$) are updated by applications, the computations are also accumulated in $ks$ and/or $kv$.
%Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
Thanks to these accumulations, there are only two $cpg$ calls in $\cpginline{bx_1 \circ bx_2}$. The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, dummy)$ where $s$ and $ks$ are corresponding to the source and the update over source. Since there is no real view here, we need a dummy same as $pg$. Then the continuation updating on this dummy should be initiated as the identity function. The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$. In the next assignment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input. It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$. After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling recursively one more time like in $\pginline{bx_1 \circ bx_2}$.

@@

In general, the beginning of a continuation should be the identity function. However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source needs to be initiated as the constant function from that source. This constant function helps to retain the discarded things in the source.

@@

The problem for $cpg$ lies in redundant computations during the evaluation. To prevent such redundant computations from occurring, we introduce an extension named $kpg$. While $cpg$ evaluates values eagerly, $kpg$ does the opposite. Every value is evaluated lazily in a computation of $kpg$. The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, lks, lkv, s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $lks$ and $lkv$ are used for lazy evaluation of actual values. The output of this function is also a 6-tuple $(ks, kv, lks, lkv, s, v)$.

Suppose that we have a source $s_0$ and a view $v_0$. The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:

@@

The beginning of accumulative functions $lks$ and $lkv$ are set as the identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$. The relationship among $kpg$, $cpg$ and $pg$ can be shown as follows:

@@

\subsection{Combination of pg and kpg: xpg}

The purpose we introduced $cpg$ and $kpg$ is to avoid redundant recursive call and keep the dropped parts from the source in a function.
On the other hand, these accumulations in $cpg$ and $kpg$ will be an overhead if they are not necessary. The problem in $\pginline{bx_1 \circ bx_2}$ is that there are two recursive calls of $\pginline{bx_1}$ and there is no problem in the recursive call of $\pginline{bx_2}$.
Therefore, we combine $pg$ and $kpg$ to take advantage of both approaches.

%The most important thing when using to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that are used to produce new source instead of using more recursive call.

%This is reflected in the first call $\kpginline{bx_1}$ and the last application in $\kpginline{bx_1 \circ bx_2}$. The second call $\kpginline{bx_2}$ just play a syntactic role while it requires many parameters. We can completely reduce this number to 2, including one for the source and one for the view as in a $pg$ call. This idea is realized by function $xpg$ as follows:

### 2020-02-21 paper/main.tex
modified the beginning of 4.4

\institute{National Institute of Informatics, Japan
  (\email{\{k\_tsushima,bach\}@nii.ac.jp}) \and
  University of Copenhagen, Denmark
  (\email{glueck@acm.org}) \and
  Peking University, China
  (\email{huzj@pku.edu.cn})}

@@

>>>>

\institute{National Institute of Informatics, Japan \break
  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
  University of Copenhagen, Denmark \break
  \email{glueck@acm.org} \and
  Peking University, China \break
  \email{huzj@pku.edu.cn}}

@@

\pagestyle{empty}
